Software Development in Startup Companies: The Green-field Startup Model

### 2013-03-03 article/main.tex
fixed comments version 0.4

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

@@

% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{figures}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
%   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
   %\graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   %\DeclareGraphicsExtensions{.eps}
%\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

@@

\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

@@

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

@@

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007  
% IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer  
% Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

@@

With a cross-sectional study we focused the research context in a limited 
time-frame to emphasize the nature of uncertainty in the development activities 
and time pressure from the market, as discussed in \cite{Maccormack2001, 
Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and 
CTOs of startups, covering a wide spectrum of thematics and iteratively 
adjusting the direction of the research according to the emerging evidences.

@@

%MUN: This is maybe only a minor concern, but it may be relevant for certain 
%reviewers: The references you use in the introduction are not very recent. 
%Look at the list of papers [1-15], most of them are published before 2005.
%I understand that you use some references to strengthen your general arguments 
%(which can refer to older publications), but your introduction should also 
%show the relevance of your research right now. This is best shown by 
%referring to recent (max 5 years old) literature. Think about if you can add 
%more recent (2010-now), well published, papers in the introduction.

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

@@

Modern entrepreneurship has been boosted by the advent of the consumer internet 
markets in the middle of the 1990s and culminated with the notorious 
\textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, 
storey1982entrepreneurship}. Several years later, with the massification of the 
internet and mobile devices, we are now assisting to an impressive proliferation 
of software ventures - metaphorically referred as the \textit{startup bubble}. 
The easy access to vast potential markets and the low cost of services 
distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}. 
Inspired by success stories, a large number of software businesses are created 
every day. However, the great majority of these companies fail within two years 
from their creation \cite{Crowne2002}.

@@

The word \textit{startup} appeared in the SE literature for the first time in 
1994 in an article written by Carmel \cite{Camel1994a} where he studied the 
time-to-completion in a young package firm. %MUN: If this is a quote 
%from Carmel, put it in quotation marks, like you did in the previous section 
%with Suttons quote. Generally, check that you are consistent how you use 
%quotations throughout the paper. 
%CG: fixed, btw  it was just to emphasize his the focus of his study. 
%MUN1: Ok. As you may have noticed, I have removed a lot of cases were you use 
%italics. If you emphasize *too* much, it it rather distracting than helpful 
%for the reader. Furthermore, it conflicts with you convention you use to 
%highlight high level categories of GSM.
He noticed how these companies were particularly innovative and successful, 
advocating for the need of more research investigating software development 
practices so as to replicate success and try to transfer it to other technology 
sectors.

@@

But when startups have no time for training and orienting activities, as 
discussed in \cite{Sutton2000}, their main focus remains on team capabilities 
instead of prescriptive processes, hiring people who can \textit{``hit the 
ground running''} \cite{Yoffie1999}. Empowering the team and focus on 
methodological attributes of the processes oriented in prototyping, 
proof-of-concepts, mock-ups and demos, to test basic functionalities, have been 
the primary priority in startups \cite{Camel1994a}. Only when they grow, formal 
methodologies arise, followed by performing quality assurance and long-term 
planning processes \cite{Yoffie1999}.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=5in]{figures/completemethodology2}
\caption{Grounded theory study process overview}
\label{fig:gt:completemethodology}
\end{figure*}
%MUN: I think you can increase the font size in the figure a bit. Try thereby
%to keep the height of the figure, expanding however the width (use horizontal
%space since it is a two column figure). Change also in figure:  - Write "Data
%Analysis" horizontally  - Change names of macro-phases to upper-case, i.e.
%Design and Execution,     Data Collection, Data Analysis, etc.  - Add section
%number to macro-phase name, i.e. 3.1 Design and Execution, 3.2     Data
%Collection  - Change "Discussion and Validation" into "Theory Validation" as
%you name it     in the text. Also, add a section (3.5) that describes shortly
%what you did in "Theory  Validation" 
%CG: fixed... fontsize cannot be increased uniformly among all the boxes, then 
%I skipped that part...
%MUN1: OK, but in the figure, the correct section number for theory validation 
%is 3.5, not 3.4

@@

Initially the boundaries of the research domain were set by means of 
non-systematic literature survey, which provided the initial keywords used 
further in the research process. Moreover it helped us in defining the research 
questions (RQs):

@@

The whole procedure has been executed simultaneously in pair on the same screen 
by the first two authors, handling conflicts by reviewing the rationale of 
decision with the third and fourth authors. When necessary we performed an 
in-depth review of the design of the research and data collected during the 
execution process.

@@

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Companies overview} 
\label{t_interviews-stats}
\centering
\begin{tabular}{|c||c||c||c||c||c|}

@@

\hline
\end{tabular}
\end{table*}

@@

%------------------------------------------------------------------------------%
\subsection{Theory Validation}
\label{rm:val}
%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e.
%the section following the GSM)
%CG: fixed
%MUN1: This description is way too long and verbose. It should be MAX 1 column, 
%and focus on what you present in Section 5. Theory Validation. Until the 
%second-last paragraph you don't talk about that at all! Remove the "theory" on 
%GT validation and what you did *during* the GT process. Keep what you need 
%to motivate
% - 5.1 Comparison of frameworks
% - 5.2 Theoretical categories and existing literature
% - 5.3 Confounding factors
% - 5.4 Static assessment of the GSM
%You need to explain why you did this and how it contributes to the validation 
%of the theory.

Presenting GT is challenging for a researcher, who must pay attention to  
structure the included level of detail, and to the way data is portrayed to 
display evidence of the emergent categories.  In order to assess our research 
and determine if the GT is sufficiently grounded, we used a systematic technique 
to assess the theory, following evaluation lists and factors presented in  
\cite{Strauss1998}.

We consider the following factors:

\par\vspace {0.4cm}
\begin{compactitem}
\item \textit{Fit}, that describes the consistency of the theory with the real 
data.
\item \textit{Understanding}, that describes the degree to which participants 
agree with the theory.
\item \textit{Generality}, that describes the level of abstraction of the 
theory, and how sufficiently it can be adopted by practitioners without losing 
its relevance.
\item \textit{Control}, that describes how well the theory enables 
practitioners to fully understand the described situation.
\end{compactitem}
\par\vspace {0.4cm}

In order to achieve a high-level of fit between the theory and the raw 
empirical data, we started GT without any assumption on how the emergent 
patterns would influence the research direction. Ideas and perceptions evolved 
during the mapping study and interviews. In addition, constant comparative 
methods, overturning of some early categories as new data came to light, 
generation and testing of interim hypotheses, and constant re-evaluation of the 
interview transcripts, ensured that researchers' bias was minimized and 
theoretical fit maintained.

To represent understandable emergent theories, concepts and categories were 
carefully developed using appropriate and self-explanatory codes. The developing 
theory was not immediately presented to the participants to prevent potential 
bias in the responses. But rather, the theoretical framework has been presented 
and described to startups practitioners not involved in the data collection, 
whose reaction was positively expressed.

From generality perspective, we developed a theory from abstract categories, 
which allowed us to create a general guide to constantly changing situations. 
Nevertheless, we took attention on maintaining a reasonable level of 
abstraction, keeping focus on their sensitizing characteristics and providing 
low-level subgroup of categories.

Finally, we provided a comprehensive set of categories with detailed 
interrelations to explain how the \textit{core-category} is achieved and the 
reasons of its consequences, in order to enable practitioners to independently 
understand and analyze situations, predict changes and their consequences, and 
to be capable of revising actions or the theory itself. Using both 
methodological tools and features provided by \textit{AtlasTI}, we explored and 
tested:

\par\vspace {0.4cm}
\begin{compactitem}
\item Each category and the strength of relations between them.
\item Hypotheses, derived from and related to the emergent theory.
\item Deviant cases to ensure robustness and general applicability.
\end{compactitem}
\par\vspace {0.4cm}

Additionally, Strauss and Corbin provided a list of questions to assist in 
determining how well the findings are grounded \cite{Strauss1998}. They 
questions are as follows:

@@

The process has been designed by the researches in different steps, explaining 
and making clear the purpose of each of them. Thus, the same processes enable 
other research to conduct the same study within similar contexts. Moreover 
\textit{comparison with the state-of-the-art} has been applied in order to 
validate the theory and enhance odds of its applicability within a wider 
time-frame. By this comparison we  highlight the areas which have been neglected 
by existing studies, providing possible directions for future studies (see 
subsection \ref{sect:theory:validation:others}). Moreover we study the 
confounding factors which could interfere with the application of the GSM model 
(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).

@@

%------------------------------------------------------------------------------%

\section{Results: Green-field startup model}
\label{res:gsm}

@@

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{SMS \cite{SMS} and GSM} 
\label{tab:an:literature_comp}
\centering
\begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}

>>>>

% Some very useful LaTeX packages include: (uncomment the ones you want to load)

@@

% *** GRAPHICS RELATED PACKAGES *** % %\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}   % declare the path(s) where your graphic files
are    \graphicspath{{figures}}   % and their extensions so you won't have to
specify these with   % every instance of \includegraphics   %
\DeclareGraphicsExtensions{.pdf,.jpeg,.png} %\else   % or other class option
(dvipsone, dvipdf, if not using dvips). graphicx   % will default to the driver
specified in the system graphics.cfg if no   % driver is specified. %
\usepackage[dvips]{graphicx}   % declare the path(s) where your graphic files
are    %\graphicspath{{./figures/}}   % and their extensions so you won't have
to specify these with   % every instance of \includegraphics
%\DeclareGraphicsExtensions{.eps} %\fi % graphicx was written by David Carlisle
and Sebastian Rahtz. It is % required if you want graphics, photos, etc.
graphicx.sty is already % installed on most LaTeX systems. The latest version
and documentation can % be obtained at:  % http://www.ctan.org/tex-
archive/macros/latex/required/graphics/ % Another good source of documentation
is "Using Imported Graphics in % LaTeX2e" by Keith Reckdahl which can be found
as epslatex.ps or % epslatex.pdf at: http://www.ctan.org/tex-archive/info/ % %
latex, and pdflatex in dvi mode, support graphics in encapsulated % postscript
(.eps) format. pdflatex in pdf mode supports graphics % in .pdf, .jpeg, .png and
.mps (metapost) formats. Users should ensure % that all non-photo figures use a
vector format (.eps, .pdf, .mps) and % not a bitmapped formats (.jpeg, .png).
IEEE frowns on bitmapped formats % which can result in "jaggedy"/blurry
rendering of lines and letters as % well as large increases in file sizes. % %
You can find documentation about the pdfTeX application at: %
http://www.tug.org/applications/pdftex

@@

% *** SPECIALIZED LIST PACKAGES *** % %\usepackage{algorithmic} %
algorithmic.sty was written by Peter Williams and Rogerio Brito. % This package
provides an algorithmic environment fo describing algorithms. % You can use the
algorithmic environment in-text or within a figure % environment to provide for
a floating algorithm. Do NOT use the algorithm % floating environment provided
by algorithm.sty (by the same authors) or % algorithm2e.sty (by Christophe
Fiorio) as IEEE does not use dedicated % algorithm float types and packages that
provide these will not provide % correct IEEE style captions. The latest version
and documentation of % algorithmic.sty can be obtained at: % http://www.ctan.org
/tex-archive/macros/latex/contrib/algorithms/ % There is also a support site at:
% http://algorithms.berlios.de/index.html % Also of interest may be the
(relatively newer and more customizable) % algorithmicx.sty package by Szasz
Janos: % http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

%\usepackage{mdwmath} %\usepackage{mdwtab} % Also highly recommended is Mark
Wooding's extremely powerful MDW tools, % especially mdwmath.sty and mdwtab.sty
which are used to format equations % and tables, respectively. The MDWtools set
is already installed on most % LaTeX systems. The lastest version and
documentation is available at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/mdwtools/

@@

\usepackage{eqparbox} % Also of notable interest is Scott Pakin's eqparbox
package for creating % (automatically sized) equal width boxes - aka "natural
width parboxes". % Available at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES *** %\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure} %\else
%\usepackage[tight,footnotesize]{subfigure} %\fi % subfigure.sty was written by
Steven Douglas Cochran. This package makes it % easy to put subfigures in your
figures. e.g., "Figure 1a and 1b". For IEEE % work, it is a good idea to load it
with the tight package option to reduce % the amount of white space around the
subfigures. Computer Society papers % use a larger font and \sffamily font for
their captions, hence the % additional options needed under compsoc mode.
subfigure.sty is already % installed on most LaTeX systems. The latest version
and documentation can % be obtained at: % http://www.ctan.org/tex-
archive/obsolete/macros/latex/contrib/subfigure/ % subfigure.sty has been
superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc %  \usepackage[caption=false]{caption} %
\usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig} %\else %
\usepackage[caption=false]{caption} %  \usepackage[font=footnotesize]{subfig}
%\fi % subfig.sty, also written by Steven Douglas Cochran, is the modern %
replacement for subfigure.sty. However, subfig.sty requires and % automatically
loads Axel Sommerfeldt's caption.sty which will override % IEEEtran.cls handling
of captions and this will result in nonIEEE style % figure/table captions. To
prevent this problem, be sure and preload % caption.sty with its "caption=false"
package option. This is will preserve % IEEEtran.cls handing of captions.
Version 1.3 (2005/06/28) and later  % (recommended due to many improvements over
1.2) of subfig.sty supports % the caption=false option directly:
%\ifCLASSOPTIONcompsoc %
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else %  \usepackage[caption=false,font=footnotesize]{subfig} %\fi % % The
latest version and documentation can be obtained at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/subfig/ % The latest version and documentation of
caption.sty can be obtained at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/caption/

@@

%\usepackage{stfloats} % stfloats.sty was written by Sigitas Tolusis. This
package gives LaTeX2e % the ability to do double column floats at the bottom of
the page as well % as the top. (e.g., "\begin{figure*}[!b]" is not normally
possible in % LaTeX2e). It also provides a command: %\fnbelowfloat % to enable
the placement of footnotes below bottom floats (the standard % LaTeX2e kernel
puts them above bottom floats). This is an invasive package % which rewrites
many portions of the LaTeX2e float routines. It may not work % with other
packages that modify the LaTeX2e float routines. The latest % version and
documentation can be obtained at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/sttools/ % Documentation is contained in the
stfloats.sty comments as well as in the % presfull.pdf file. Do not use the
stfloats baselinefloat ability as IEEE % does not allow \baselineskip to
stretch. Authors submitting work to the % IEEE should note that IEEE rarely uses
double column equations and % that authors should try to avoid such use. Do not
be tempted to use the % cuted.sty or midfloat.sty packages (also by Sigitas
Tolusis) as IEEE does % not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff %  \usepackage[nomarkers]{endfloat} %
\let\MYoriglatexcaption\caption %
\renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}} %\fi %
endfloat.sty was written by James Darrell McCauley and Jeff Goldberg. % This
package may be useful when used in conjunction with IEEEtran.cls' % captionsoff
option. Some IEEE journals/societies require that submissions % have lists of
figures/tables at the end of the paper and that % figures/tables without any
captions are placed on a page by themselves at % the end of the document. If
needed, the draftcls IEEEtran class option or % \CLASSINPUTbaselinestretch
interface can be used to increase the line % spacing as well. Be sure and use
the nomarkers option of endfloat to % prevent endfloat from "marking" where the
figures would have been placed % in the text. The two hack lines of code above
are a slight modification of % that suggested by in the endfloat docs (section
8.3.1) to ensure that % the full captions always appear in the list of
figures/tables - even if % the user used the short optional argument of
\caption[]{}. % IEEE papers do not typically make use of \caption[]'s optional
argument, % so this should not be an issue. A similar trick can be used to
disable % captions of packages such as subfig.sty that lack options to turn off
% the subcaptions: % For subfig.sty: % \let\MYorigsubfloat\subfloat %
\renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}} % For subfigure.sty:
% \let\MYorigsubfigure\subfigure %
\renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}} % However, the
above trick will not work if both optional arguments of % the \subfloat/subfig
command are used. Furthermore, there needs to be a % description of each
subfigure *somewhere* and endfloat does not add % subfigure captions to its list
of figures. Thus, the best approach is to % avoid the use of subfigure captions
(many IEEE journals avoid them anyway) % and instead reference/explain all the
subfigures within the main caption. % The latest version of endfloat.sty and its
documentation can obtained at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/endfloat/ % % The IEEEtran
\ifCLASSOPTIONcaptionsoff conditional can also be used % later in the document,
say, to conditionally put the References on a  % page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES *** % \usepackage{url} % url.sty was
written by Donald Arseneau. It provides better support for % handling and
breaking URLs. url.sty is already installed on most LaTeX % systems. The latest
version can be obtained at: % http://www.ctan.org/tex-
archive/macros/latex/contrib/misc/ % Read the url.sty source comments for usage
information. Basically, % \url{my_url_here}.

@@

% The publisher's ID mark at the bottom of the page is less important with %
Computer Society journal papers as those publications place the marks % outside
of the main text columns and, therefore, unlike regular IEEE % journals, the
available text space is not reduced by their presence. % If you want to put a
publisher's ID mark on the page you can do it like % this: %\IEEEpubid{0000--
0000/00\$00.00~\copyright~2007 IEEE} % or like this to get the Computer Society
new two part style. %\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--
0000/00/\$00.00~\copyright~2007   % IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer   %
Society\hfill}} % Remember, if you use this you must call \IEEEpubidadjcol in
the second % column for its text to clear the IEEEpubid mark (Computer Society
jorunal % papers don't need this extra clearance.)

@@

With a cross-sectional study we focused the research context in a limited  time-
frame to emphasize the nature of uncertainty in the development activities  and
time pressure from the market, as discussed in \cite{Maccormack2001,
Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and
CTOs of startups, covering a wide spectrum of thematics and iteratively
adjusting the direction of the research according to the emerging evidences.

@@

%MUN: This is maybe only a minor concern, but it may be relevant for certain
%reviewers: The references you use in the introduction are not very recent.
%Look at the list of papers [1-15], most of them are published before 2005. I
%understand that you use some references to strengthen your general arguments
%(which can refer to older publications), but your introduction should also
%show the relevance of your research right now. This is best shown by  referring
%to recent (max 5 years old) literature. Think about if you can add  more recent
%(2010-now), well published, papers in the introduction.

% You must have at least 2 lines in the paragraph with the drop letter (should
% never be an issue)

@@

Modern entrepreneurship has been boosted by the advent of the consumer internet
markets in the middle of the 1990s and culminated with the notorious  \textit
{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126,
storey1982entrepreneurship}. Several years later, with the massification of the
internet and mobile devices, we are now assisting to an impressive proliferation
of software ventures - metaphorically referred as the \textit{startup bubble}.
The easy access to vast potential markets and the low cost of services
distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}.
Inspired by success stories, a large number of software businesses are created
every day. However, the great majority of these companies fail within two years
from their creation \cite{Crowne2002}.

@@

The word \textit{startup} appeared in the SE literature for the first time in
1994 in an article written by Carmel \cite{Camel1994a} where he studied the
time-to-completion in a young package firm. %MUN: If this is a quote  %from
Carmel, put it in quotation marks, like you did in the previous section  %with
Suttons quote. Generally, check that you are consistent how you use  %quotations
throughout the paper.  %CG: fixed, btw  it was just to emphasize his the focus
of his study.  %MUN1: Ok. As you may have noticed, I have removed a lot of cases
were you use  %italics. If you emphasize *too* much, it it rather distracting
than helpful  %for the reader. Furthermore, it conflicts with you convention you
use to  %highlight high level categories of GSM. CG: ok. He noticed how these
companies were particularly innovative and successful,  advocating for the need
of more research investigating software development  practices so as to
replicate success and try to transfer it to other technology  sectors.

@@

But when startups have no time for training and orienting activities, as
discussed in \cite{Sutton2000}, their main focus remains on team capabilities
instead of prescriptive processes, hiring people who can \textit{``hit the
ground running''} \cite{Yoffie1999}. Empowering the team and focus on
methodological attributes of the processes oriented in prototyping,  proof-of-
concepts, mock-ups and demos, to test basic functionalities, have been  the
primary priority in startups \cite{Camel1994a}. Only when they grow, formal
methodologies arise, followed by performing quality assurance and long-term
planning processes \cite{Yoffie1999}.

@@

\begin{figure*}[!t] \centering
\includegraphics[width=5in]{figures/completemethodology2} \caption{Grounded
theory study process overview} \label{fig:gt:completemethodology} \end{figure*}
%MUN: I think you can increase the font size in the figure a bit. Try thereby
%to keep the height of the figure, expanding however the width (use horizontal
%space since it is a two column figure). Change also in figure:  - Write "Data
%Analysis" horizontally  - Change names of macro-phases to upper-case, i.e.
%Design and Execution,     Data Collection, Data Analysis, etc.  - Add section
%number to macro-phase name, i.e. 3.1 Design and Execution, 3.2     Data
%Collection  - Change "Discussion and Validation" into "Theory Validation" as
%you name it     in the text. Also, add a section (3.5) that describes shortly
%what you did in "Theory  Validation"  %CG: fixed... fontsize cannot be
increased uniformly among all the boxes, then  %I skipped that part... %MUN1:
OK, but in the figure, the correct section number for theory validation  %is
3.5, not 3.4. CG:fixed

@@

Initially the boundaries of the research domain were set by means of  non-
systematic literature survey, which provided the initial keywords used  further
in the research process. Moreover it helped us in defining the research
questions (RQs):

@@

The whole procedure has been executed simultaneously in pair on the same screen
by the first two authors, handling conflicts by reviewing the rationale of
decision with the third and fourth authors. When necessary we performed an  in-
depth review of the design of the research and data collected during the
execution process.

@@

\begin{table*}[!t] renewcommand{arraystretch}{1.3} caption{Companies overview}
\label{t_interviews-stats} centering begin{tabular}{|c||c||c||c||c||c|}

@@

\hline end{tabular} end{table*}

@@

%------------------------------------------------------------------------------%
\subsection{Theory Validation} \label{rm:val} %MUN: put here what I wrote in the
comment in Section "Theory validation" (i.e. %the section following the GSM)
%CG: fixed %MUN1: This description is way too long and verbose. It should be MAX
1 column,  %and focus on what you present in Section 5. Theory Validation. Until
the  %second-last paragraph you don't talk about that at all! Remove the
"theory" on  %GT validation and what you did *during* the GT process. Keep what
you need  %to motivate % - 5.1 Comparison of frameworks % - 5.2 Theoretical
categories and existing literature % - 5.3 Confounding factors % - 5.4 Static
assessment of the GSM %You need to explain why you did this and how it
contributes to the validation  %of the theory. CG: fixed, removing the initial
verbose part.

Presenting GT is challenging for a researcher, who must pay attention to
structure the included level of detail, and to the way data is portrayed to
display evidence of the emergent categories.  In order to assess our research
and determine if the GT is sufficiently grounded, we used a systematic technique
to assess the theory.

Strauss and Corbin provided a list of questions to assist in  determining how
well the findings are grounded \cite{Strauss1998}. They  questions are as
follows:

@@

The process has been designed by the researches in different steps, explaining
and making clear the purpose of each of them. Thus, the same processes enable
other research to conduct the same study within similar contexts. Moreover
\textit{comparison with the state-of-the-art} has been applied in order to
validate the theory and enhance odds of its applicability within a wider  time-
frame. By this comparison we  highlight the areas which have been neglected  by
existing studies, providing possible directions for future studies (see
subsection \ref{sect:theory:validation:others}). Moreover we study the
confounding factors which could interfere with the application of the GSM model
(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).

@@

\section{Results: Green-field startup model} label{res:gsm}

@@

\begin{table*}[!t] renewcommand{arraystretch}{1.3} caption{SMS cite{SMS} and
\GSM}  label{tab:an:literature_comp} centering
\begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}
